---
title: "Test title"
output: 
  html_document:
    mathjax: "default"
---



```{r init, include = FALSE}
library(ggplot2)
library(patchwork)
```

As shown in the documentation for this project, it is possible to have the same summary statistics but different underlying household configurations.

# Definitions and Notation

Let:

- \(P\) = **Total population size**
- \(H\) = **Number of households** 
- \(M_p\) = **Person-level mean household size**

    For every person \( j \) in a population \( P \) living in a household of \( m_j \) members, the **person-level** mean household size is a simple average across person-level observations:
    
    \begin{equation}
    \label{eq:1}
    \tag{1}
    M_p = \frac{\sum_{j = 1}^H m_j^2}{P}
    \end{equation}

- \(M_h\) = **Household-level mean household size**

   The **household-level** mean household size is a simple average across households as individual observations:
    
    \begin{align}
    \label{eq:2}
    \tag{2}
    M_h = \frac{P}{H}
    \end{align}

---

# Context

Equation \eqref{eq:2} shows that household-level mean household size and population is adequate to deduce the number of households. However, person-level household size does not lend itself to this equivalency: the same person-level mean and population can be consistent with multiple values for the number of households. 

Figure 1.2 illustrates this point in a hypothetical population with \( P = 6 \) and \( M_p = 3 \). In Scenario 1, average person-level household size is \( \frac{3^2 + 3^2}{6} = 3 \) and there are 2 households. In Scenario 2, average person-level household size is \( \frac{1^2 + 1^2 + 4^2}{6} = 3 \) and there are 3 households. Thus, population and person-level household size alone do not singularly determine the number of households.

**Figure 1.2**

```{r fig01.2, echo=FALSE, out.width="100%"}
# Data for Scenario 1
hh_data1 <- data.frame(
  hh_id = rep(c("Household A", "Household B"), times = c(3, 3)),
  x = c(1, 2, 1.5, 1, 2, 1.5),
  y = c(1, 1, 2, 1, 1, 2)
)

# Data for Scenario 2
hh_data2 <- data.frame(
  hh_id = rep(c("Household A", "Household B", "Household C"), times = c(1, 1, 4)),
  x = c(1.5, 1.5, 1, 1, 2, 2),
  y = c(1.5, 1.5, 1, 2, 1, 2)
)

# Plot for Scenario 1
plot1 <- ggplot(hh_data1, aes(x = x, y = y)) +
  geom_point(shape = 21, size = 4, color = "black", stroke = 1.2) +
  theme_void() +
  theme(
    plot.margin = margin(1, 1, 1, 1),
    plot.background = element_rect(color = "black", fill = NA, linewidth = 0.5)
  ) +
  facet_wrap(~hh_id, nrow = 1) +
  coord_fixed(ratio = 1) +
  xlim(0, 3) +
  ylim(0, 3) +
  ggtitle("Scenario 1")

# Plot for Scenario 2
plot2 <- ggplot(hh_data2, aes(x = x, y = y)) +
  geom_point(shape = 21, size = 4, color = "black", stroke = 1.2) +
  theme_void() +
  theme(
    plot.margin = margin(1, 1, 1, 1),
    plot.background = element_rect(color = "black", fill = NA, linewidth = 0.5)
  ) +
  facet_wrap(~hh_id, nrow = 1) +
  coord_fixed(ratio = 1) +
  xlim(0, 3) +
  ylim(0, 3) +
  ggtitle("Scenario 2")
```

```{r, fig.height=3, out.width="100%", echo = FALSE}
# Combine plots using patchwork
combined_plot <- plot1 + plot2 +
  plot_layout(widths = c(1, 1)) +
  plot_annotation(
    caption = expression(paste("Circles represent individual people. In both scenarios, P = 6 and ", M[p], " = 3, yet the number of households differs."))
  )

# Display the plot
print(combined_plot)

```

---

# Assertion

There is a range of possible number of households \(H\) consistent with a known population size \(P\) and person-level average household size \(M_p\). The range on \(H\) is bounded by a minumum of 

\begin{align}
H_{min} = \frac{P}{M_p}
\end{align}

and a maximum of

[FILL IN HERE]

---

# Lemma 1: Minimum Number of Households  

The minimum number of households consistent with a population size \(P\) and person-level average household size \(M_p\) is:
\begin{align}
H_{\min} = \frac{P}{M_p}
\end{align}

## Proof

Suppose the population $P$ is distributed into $H$ households with the measured household size in the $j$th household equal to $m_j$. As defined above, **person-level** mean household size is:
    
\begin{equation}
M_p = \frac{1}{P}\sum_{j = 1}^H m_j^2
\end{equation}

Any configuration of $P$ people among households can be achieved by incrementally moving one individual at a time from one household to another. Forming a new household can be achieved by moving an individual from an existing household to household $H+1$ (technically out of the original range of number of households, but the total number of households in any configuration can be regarded as H households with a positive number of residents and an infinite number of 0-person households). For example, one can move from Scenario 1 to Scenario 2 in Figure 1.2 through the following steps:

1. Move one individual from Household A to a (newly populated) Household C
2. Move one of the remaining 2 individuals in Household A to Household C
3. Move one individual from Household B to Household C
4. Move one of the remaining 2 individuals in Household B to Household C
5. The configuration now matches Scenario 2.

Note that in our notation, we

Suppose we begin with an arbitrary configuration of $P$ individuals in $H$ households, with each household $m_j$ containing some whole number of members: $m_1, m_2, \ldots, m_{H-1}, m_H$. We move one individual from household $j = k$ to household $j = k+1$. Then mean household size changes from:

\begin{equation}
M_p =\frac{1}{P} \bigg( \sum_{j = 1}^{k-1} m_j^2 + m_k^2 + m_{k+1}^2 +  \sum_{j = k+2}^{H+1}m_j^2 \bigg)
\end{equation}

to 

\begin{equation}
M_p' =\frac{1}{P} \bigg( \sum_{j = 1}^{k-1} m_j'^2 + m_k'^2 + m_{k+1}'^2 + \sum_{j = k+2}^{H+1}m_j'^2 \bigg)
\end{equation}

Person-level household size increases if $M_p' - M_p > 0$, remains the same if $M_p' - M_p = 0$, and shrinks if $M_p' - M_p < 0$. Since all household sizes $m_j = m_j'$ are unaffected by the shift, aside from the special case of $j = k$ and $j = k+1$, the remaining terms cancel out:

\begin{align}
M_p' - M_p & = \frac{1}{P} \bigg( \sum_{j = 1}^{k-1} m_j^2 + m_k^2 + m_{k+1}^2 +  \sum_{j = k+2}^{H+1}m_j^2 \bigg) - \frac{1}{P} \bigg( \sum_{j = 1}^{k-1} m_j'^2 + m_k'^2 + m_{k+1}'^2 + \sum_{j = k+2}^{H+1}m_j'^2 \bigg) \\
& = \frac{1}{P} \bigg( (m_k^2 + m_{k+1}^2) - (m_k'^2 + m_{k+1}'^2) \bigg) \\
\end{align}
Since only households $j = k$ and $j = k+1$ are affected by this change, the remaining sums can be regarded as equivalent and cancelled out.


...

this section of the proof ends where I show that there's no possible way to "dissolve" a household and add its contents to other households without pushing the average persn-level household size above $m$.

Case B:
Then I illustrate the same concept even when $i$ is not an integer by creating the "closest square household configuration" where I essentially show that, for example, if m = 2.75, you can configure the population into groups of the nearest integer-size households (2- and 3-person housholds) such that dissolving the contents of any one household and adding it to other households ill push the average person-level household size above $m$.

Thus, the configurations described above give the minimum number of households, $i = p/m$. And, in fact, when $i$ is not an integer, as in Case B, then the true minimum is given by rounding $i$ up to the nearest integer.

---

# Lemma 2: Maximum Number of Households  
**Lemma:**  
The maximum number of households consistent with a population size \(P\) and person-level average household size \(M_p\) is:
\begin{align}
TBD
\end{align}

**Proof:**  
In this section, I show that the sparsest configuraiton of households is given in a case where you produce some number of 1-person households and then throw the rest of the population into a remaining large households.

---
# conclusion

Reiterate the minima and maxima, take it back to the example at the begining and show that the two configurations I listed are actually the two special cases of minima and maxima.
