---
title: "Test title"
output: 
  html_document:
    mathjax: "default"
---



```{r init, include = FALSE}
library(ggplot2)
library(patchwork)
```

As shown in the documentation for this project, it is possible to have the same summary statistics but different underlying household configurations.

# Definitions and Notation

Let:

- \(P\) = **Total population size**
- \(H\) = **Number of households** 
- \(M_p\) = **Person-level mean household size**

    For every person \( j \) in a population \( P \) living in a household of \( m_j \) members, the **person-level** mean household size is a simple average across person-level observations:
    
    \begin{equation}
    \label{eq:1}
    \tag{1}
    M_p = \frac{\sum_{j = 1}^P m_j}{P}
    \end{equation}

- \(M_h\) = **Household-level mean household size**

   The **household-level** mean household size is a simple average across households as individual observations:
    
    \begin{align}
    \label{eq:2}
    \tag{2}
    M_h = \frac{P}{H}
    \end{align}

---

# Context

Equation \eqref{eq:2} shows that household-level mean household size and population is adequate to deduce the number of households. However, person-level household size does not lend itself to this equivalency: the same person-level mean and population can be consistent with multiple values for the number of households. 

Figure 1.2 illustrates this point in a hypothetical population with \( P = 6 \) and \( M_p = 3 \). In scenario 1, average person-level household size is \( \frac{3+3+3 + 3+3+3}{6} = 3 \) and there are 3 households. In scenario 2, average person-level household size is \( \frac{1 + 1 + 4+4+4+4}{6} = 3 \) and there are 4 households. Thus, population and person-level household size alone do not singularly determine the number of households.

**Figure 1.2**

```{r fig01.2, echo=FALSE, out.width="100%"}
# Data for Scenario 1
hh_data1 <- data.frame(
  hh_id = rep(c("Household A", "Household B"), times = c(3, 3)),
  x = c(1, 2, 1.5, 1, 2, 1.5),
  y = c(1, 1, 2, 1, 1, 2)
)

# Data for Scenario 2
hh_data2 <- data.frame(
  hh_id = rep(c("Household A", "Household B", "Household C"), times = c(1, 1, 4)),
  x = c(1.5, 1.5, 1, 1, 2, 2),
  y = c(1.5, 1.5, 1, 2, 1, 2)
)

# Plot for Scenario 1
plot1 <- ggplot(hh_data1, aes(x = x, y = y)) +
  geom_point(shape = 21, size = 4, color = "black", stroke = 1.2) +
  theme_void() +
  theme(
    plot.margin = margin(1, 1, 1, 1),
    plot.background = element_rect(color = "black", fill = NA, linewidth = 0.5)
  ) +
  facet_wrap(~hh_id, nrow = 1) +
  coord_fixed(ratio = 1) +
  xlim(0, 3) +
  ylim(0, 3) +
  ggtitle("Scenario 1")

# Plot for Scenario 2
plot2 <- ggplot(hh_data2, aes(x = x, y = y)) +
  geom_point(shape = 21, size = 4, color = "black", stroke = 1.2) +
  theme_void() +
  theme(
    plot.margin = margin(1, 1, 1, 1),
    plot.background = element_rect(color = "black", fill = NA, linewidth = 0.5)
  ) +
  facet_wrap(~hh_id, nrow = 1) +
  coord_fixed(ratio = 1) +
  xlim(0, 3) +
  ylim(0, 3) +
  ggtitle("Scenario 2")
```

```{r, fig.height=3, out.width="100%"}
# Combine plots using patchwork
combined_plot <- plot1 + plot2 +
  plot_layout(widths = c(1, 1))

# Display the plot
print(combined_plot)

```

---

# Goal

We seek to determine the possible range for the number of households \(H\) consistent with a known population size \(P\) and person-level average household size \(M_p\). Specifically, we will establish both a minimum and maximum bound for \(H\). Given population size and average person-level household size, there is a minimum number of households given by

\begin{align}
H_{min} = \frac{P}{\text{(Mean HH Size)}_{\text{person}}}
\end{align}

and a max given by

[FILL IN HERE]

---

# Lemma 1: Minimum Number of Households  
**Lemma:**  
The minimum number of households consistent with a population size \(P\) and person-level average household size \(M_p\) is:
\begin{align}
H_{\min} = \frac{P}{M_p}
\end{align}

**Proof:**  

Case A:
Consider a specific case where $P = p$ and $M_p = m$, which happens to be a factor of $p$: in other words, there exists a positive integer $i$ such that $m * i = p$. Let us imagine a town configured in this fashion. One (but potentially not the only) configuration consistent with these statistics: $M_p = m$, $P = p$, is that the town has exactly $i$ households, each containing exactly $\frac{p}{i} = m$ members:

[schematic showing this in an example where P = 30, m = 3].

...

this section of the proof ends where I show that there's no possible way to "dissolve" a household and add its contents to other households without pushing the average persn-level household size above $m$.

Case B:
Then I illustrate the same concept even when $i$ is not an integer by creating the "closest square household configuration" where I essentially show that, for example, if m = 2.75, you can configure the population into groups of the nearest integer-size households (2- and 3-person housholds) such that dissolving the contents of any one household and adding it to other households ill push the average person-level household size above $m$.

Thus, the configurations described above give the minimum number of households, $i = p/m$. And, in fact, when $i$ is not an integer, as in Case B, then the true minimum is given by rounding $i$ up to the nearest integer.

---

# Lemma 2: Maximum Number of Households  
**Lemma:**  
The maximum number of households consistent with a population size \(P\) and person-level average household size \(M_p\) is:
\begin{align}
TBD
\end{align}

**Proof:**  
In this section, I show that the sparsest configuraiton of households is given in a case where you produce some number of 1-person households and then throw the rest of the population into a remaining large households.

---
# conclusion

Reiterate the minima and maxima, take it back to the example at the begining and show that the two configurations I listed are actually the two special cases of minima and maxima.
